---
title: "StarSchema"
author: "Andrew Lowe"
date: "8 January 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Read in the data (and drop the numeric data to save RAM):
```{r}
require(data.table)

StarSchema <- fread("C:/Work/Projects/Bayer/ws/Tableau/_work/output/StarSchema_CAT_PT_OVR.csv", 
                    select = c(2, 46, 54, 58, 59, 197), # Just these columns; we don't need the rest
                    showProgress = T)
```

```{r}
StarSchema$IS_CATEGORY_MAPPED <- factor(StarSchema$IS_CATEGORY_MAPPED)
StarSchema$MARKET_LEVEL <- factor(StarSchema$MARKET_LEVEL)
StarSchema$SOURCE_CODE <- factor(StarSchema$SOURCE_CODE)
StarSchema$Month <- factor(StarSchema$Month == "")
```

```{r}
table(StarSchema$IS_CATEGORY_MAPPED)/sum(table(StarSchema$IS_CATEGORY_MAPPED))*100
table(StarSchema$MARKET_LEVEL)/sum(table(StarSchema$MARKET_LEVEL))*100
table(StarSchema$SOURCE_CODE)/sum(table(StarSchema$SOURCE_CODE))*100
table(StarSchema$Month)/sum(table(StarSchema$Month))*100
```

```{r}
pie(table(StarSchema$IS_CATEGORY_MAPPED))
pie(table(StarSchema$MARKET_LEVEL))
pie(table(StarSchema$SOURCE_CODE))
pie(table(StarSchema$Month))
```

```
```{r}
require(digest)
dup <- duplicated(lapply(StarSchema, function(x) digest(toupper(x))))
print(dup)
StarSchema[,names(StarSchema)[dup]]
```


Make a collection of the unique items in each column:
```{r}
dat <- sapply(StarSchema, unique)
```

Define the character strings that define the market segments observed in the *Market by Brands* sheets:
```{r}
 # Note trailing spaces
segments <- c("ANTISEPTIC\ ",
              "INTIMATE HYGIENE", 
              "OTHER GYNAECOLOGICAL PREPARATIONS", 
              "Intimate Hygiene - Mycosis",
              "Calming and sleeping products",
              "Skin Repair",
              "Laxative",
              "Allergy",
              "GERD\ ", # Note trailing space
              "Suncare"
)
```

Loop over each market segment and search for it in the collection of unique items collected from the columns:
```{r}
results <- lapply(segments, function(segment) { # Loop over all defined segments
  unlist( # Collect the results
    # Search for this segment in each column:
    lapply(dat, grep, pattern = segment, ignore.case = TRUE, value = TRUE)
  )
})
names(results) <- segments
```

Print the results:
```{r}
print(results)
```

Which segment categories are missing; that is, there is no data available to enable categorisation into the segments that are present in the *Market by Brand* sheet? 
```{r}
# Is the segment missing, i.e., we couldn't find any data for it:
missing <- sapply(results, length) == 0L
print(missing[missing]) # These are missing
```

```{r}
# install.packages("benford.analysis")
require(benford.analysis)
bfd.cp <- benford(StarSchema$Sellout_TurnoverBRG[StarSchema$PHARMA_GROUP == "BAYER CC"], discrete = T, number.of.digits = 1) #generates benford object
bfd.cp #prints
plot(bfd.cp) #plots
```

```{r}
bfd.cp <- benford(StarSchema$Sellout_Units[StarSchema$PHARMA_GROUP == "BAYER CC"], number.of.digits = 1) #generates benford object
bfd.cp #prints
plot(bfd.cp) #plots
```